---
title: "Closed-Form Methodology for Correlated Group Sequential Designs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{closed-form-methodology}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(CorrSurvGSD)
```

## Scope

- `CorrSurvGSD` evaluates correlated PFS and OS group sequential designs under
the Fleischer model.
- The implementation follows the developer specification in `code/Blueprint.md`.
- This vignette summarizes the user-facing statistical workflow and the
closed-form quantities returned by the package.
- The pipeline returns theoretical results and can optionally append Monte
Carlo results without changing the theoretical calculations.

## Pipeline example

The public interface is `run_pipeline()`. It validates the design, derives the
calendar cutoffs and moments, constructs the joint correlation matrix, and
calculates power.

```{r pipeline-example}
state <- run_pipeline(
  n_T = 100,
  n_C = 100,
  d_PFS_vec = c(50),
  v_vec = c(200 / 12),
  median_PFS_C = log(2) / 0.15,
  median_OS_C = log(2) / 0.05,
  median_PFS_T = log(2) / 0.15,
  median_OS_T = log(2) / 0.05,
  alpha_spending_PFS = "OF",
  alpha_spending_OS = "Pocock",
  efficacy_looks = list(PFS = 1, OS = 1),
  futility_looks = list(PFS = integer(0), OS = integer(0))
)

state$theoretical_results$joint_correlation_matrix
```

## Model inputs and pipeline state

- The design requires treatment and control sample sizes, accrual rates, target
PFS event counts, control PFS and OS medians, and either treatment medians or
endpoint-specific hazard ratios. The unused treatment input pair defaults to
`NULL`.
- A named `efficacy_looks` list must be supplied to specify the efficacy
analyses for each endpoint.
- A named `futility_looks` list must be supplied to specify endpoint-specific
futility analyses before the final look. Use empty vectors when futility is not
used.
- `hierarchy_order` specifies whether PFS or OS is the primary gatekeeping
endpoint.
- The returned `trial_state` stores validated design quantities in
`state$design`, numerical settings in `state$options`, and theoretical
results in `state$theoretical_results`.

## Module 0: input validation and derived design quantities

- The total sample size is `n = n_T + n_C`. Piecewise accrual rates are
converted to segment durations and enrollment probabilities.
- Control-arm medians are converted to latent Fleischer hazards.

\begin{equation*}
\Lambda_C = \frac{\log(2)}{\operatorname{median}_{PFS,C}}, \qquad
\lambda_{2,C} = \frac{\log(2)}{\operatorname{median}_{OS,C}}, \qquad
\lambda_{1,C} = \Lambda_C - \lambda_{2,C}.
\end{equation*}

- Treatment hazards are derived either from treatment medians or from
`HR_PFS` and `HR_OS`. The validator rejects mixed or incomplete treatment
effect specifications.
- Alpha spending supports `"OF"`, `"Pocock"`, and `"HSD"`. HSD requires a
finite endpoint-specific gamma parameter.
- Integration and simulation seeds are stored separately so numerical
integration and Monte Carlo results can be reproduced independently.

## Module 1: calendar cutoffs

- For each PFS event look, the pipeline solves for the expected calendar
cutoff `A_l`.

\begin{equation*}
n_T m_{\delta_l^{PFS}}^{(T)}(A_l) +
n_C m_{\delta_l^{PFS}}^{(C)}(A_l) = d_l^{PFS}.
\end{equation*}

- The monotone event-count equation is solved numerically using `tol`.
The resulting `A_vec` is saved in `state$design`.

## Module 2: per-subject moments

- For a subject enrolled at time `U`, define `B = min(A, R)` and integrate
over the piecewise-uniform enrollment distribution.
- For endpoint hazard `mu`, the single-look event and truncated-time means
are obtained from the closed-form integrals `I_1`, `I_4`, and `I_5`.

\begin{aligned*}
m_\delta &= I_1(A,B,b,\mu) - I_4(A,B,b,\mu), \\
m_{\tilde{T}} &=
\frac{I_1(A,B,b,\mu)-I_4(A,B,b,\mu)}{\mu}, \\
m_{\tilde{T}^2} &=
\frac{2}{\mu^2}\left[I_1-I_4-\mu I_5\right], \\
m_{\delta\tilde{T}} &=
\frac{I_1-I_4}{\mu} - I_5.
\end{aligned*}

- PFS/OS cross-look products use the ordering of `A_1` and `A_2` and the
Fleischer latent hazards. The resulting single-look and cross-look moments
are stored in `state$moments`.
- Expected OS event counts are derived from the single-look means and saved
as `state$design$d_OS_vec`.

## Module 3: joint correlation matrix

- The log hazard-ratio variances and PFS/OS cross-endpoint covariances are
calculated with the Delta method.

\begin{equation*}
\operatorname{Var}(\hat{\theta}_l^k) =
\sum_{g \in \{T,C\}} \frac{1}{n_g}
\nabla g^\top \Sigma_{k,l}^{(g)} \nabla g.
\end{equation*}

- Within an endpoint, canonical group-sequential information gives
`rho(l_1, l_2) = sqrt(d_min / d_max)`.
- Across endpoints, the covariance is standardized by the corresponding
marginal variances. The combined `(2L) x (2L)` matrix is stored with PFS
looks first and OS looks second.

## Module 4: boundaries and theoretical power

- The mean drift at look `l` for endpoint `k` is

\begin{equation*}
\theta_l^k = -\log(HR_k)\sqrt{d_l^k\frac{r}{(1+r)^2}},
\qquad r = \frac{n_T}{n_C}.
\end{equation*}

- Efficacy boundaries are calculated from the selected alpha-spending family.
Configured futility boundaries are calculated from the corresponding
futility hazard ratios.
- The primary Z-scale boundary is retained in `state$design$boundary`.
Finite boundaries are also converted to HR and one-sided p scales.

\begin{equation*}
s_l = \sqrt{d_l\frac{r}{(1+r)^2}}, \qquad
HR_l = \exp\left(-\frac{Z_l}{s_l}\right), \qquad
p_l = 1 - \Phi(Z_l).
\end{equation*}

- `state$design$boundary_HR` and `state$design$boundary_p` use `NA_real_`
for unconfigured boundaries, while Z-scale `-Inf` and `Inf` remain available
for power and simulation logic.
- `marginal_power` is calculated from first efficacy-crossing regions using
multivariate normal integration and `integration_seed`.
- Joint power follows `hierarchy_order`: the primary endpoint must cross its
efficacy boundary before the secondary endpoint is evaluated. The mutually
exclusive crossing probabilities are stored in `joint_power_matrix` and sum
to `joint_power`.

## Module 5: optional simulation

- When `simulation = TRUE`, the package generates trial-level Fleischer event
times, enrollment times, event counts, person-time, and Z statistics.
- Simulation derives each calendar cutoff from the simulated PFS event times
and calculates OS statistics from the simulated OS event times. It does not
modify the closed-form results.
- Invalid replicates with zero events or zero person-time are regenerated up
to the internal attempt limit.
- Empirical means, correlations, marginal power, and joint power are stored in
`state$empirical_results`. The simulation seed remains in
`state$options$simulation_seed`.

## Multi-look gatekeeping example

```{r multi-look-example}
multi_state <- run_pipeline(
  n_T = 100,
  n_C = 100,
  d_PFS_vec = c(50, 100),
  v_vec = 200 / 12,
  median_PFS_C = log(2) / 0.15,
  median_OS_C = log(2) / 0.05,
  HR_PFS = 0.8,
  HR_OS = 0.8,
  alpha_spending_PFS = "OF",
  alpha_spending_OS = "Pocock",
  efficacy_looks = list(PFS = c(1, 2), OS = c(1, 2)),
  futility_looks = list(PFS = integer(0), OS = integer(0)),
  hierarchy_order = c(primary = "PFS", secondary = "OS")
)

multi_state$theoretical_results$joint_power
```

## Optional simulation example

```{r simulation-example, eval = FALSE}
simulation_state <- run_pipeline(
  n_T = 100,
  n_C = 100,
  d_PFS_vec = c(50, 100),
  v_vec = 200 / 12,
  median_PFS_C = log(2) / 0.15,
  median_OS_C = log(2) / 0.05,
  HR_PFS = 0.8,
  HR_OS = 0.8,
  alpha_spending_PFS = "OF",
  alpha_spending_OS = "Pocock",
  efficacy_looks = list(PFS = c(1, 2), OS = c(1, 2)),
  futility_looks = list(PFS = integer(0), OS = integer(0)),
  simulation = TRUE,
  simulation_seed = 777,
  n_sim = 500
)

simulation_state$empirical_results$joint_power
```

## Reproducibility

- Set `integration_seed` to control numerical multivariate-normal integration.
- Set `simulation_seed` and `simulation = TRUE` to obtain reproducible Monte
Carlo summaries.
- The returned `trial_state` preserves the validated design, options, and
theoretical results for downstream reporting.
