---
title: "Closed-Form Methodology for Correlated Group Sequential Designs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{closed-form-methodology}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(CorrSurvGSD)
```

## Scope

- `CorrSurvGSD` evaluates correlated PFS and OS group sequential designs under
the Fleischer model.
- The implementation follows the modules in `../Blueprint.md`.
- The pipeline returns theoretical results and can optionally append Monte
Carlo results without changing the theoretical calculations.

## Pipeline example

The public interface is `run_pipeline()`. It validates the design, derives the
calendar cutoffs and moments, constructs the joint correlation matrix, and
calculates power.

```{r pipeline-example}
state <- run_pipeline(
  n_T = 100,
  n_C = 100,
  d_PFS_vec = c(50),
  v_vec = c(200 / 12),
  median_PFS_C = log(2) / 0.15,
  median_OS_C = log(2) / 0.05,
  median_PFS_T = log(2) / 0.15,
  median_OS_T = log(2) / 0.05,
  alpha_spending_PFS = "OF",
  alpha_spending_OS = "Pocock"
)

state$theoretical_results$joint_correlation_matrix
```

## Reproducibility

- Set `integration_seed` to control numerical multivariate-normal integration.
- Set `simulation_seed` and `simulation = TRUE` to obtain reproducible Monte
Carlo summaries.
- The returned `trial_state` preserves the validated design, options, and
theoretical results for downstream reporting.
